---
title: "Algorithm Comparison"
output: html_document
params:
  formatted1: NA 
  formatted1_pars: NA
  formatted2: NA 
  formatted2_pars: NA
  formatted3: NA 
  formatted3_pars: NA
  approxmapR_results: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Meta Information

Algorithms being compared are `r formatted1_pars$algo` and `r formatted2_pars$algo`.
The parameters set for each algorithm are as follows:

* `r formatted1_pars$algo`: k = `r formatted1_pars$k` with the `r formatted1_pars$pattern` filter pattern being used at the threshold of `r formatted1_pars$threshold * 100`%.
* `r formatted2_pars$algo`: k = `r formatted2_pars$k` with the `r formatted2_pars$pattern` filter pattern being used at the threshold of `r formatted2_pars$threshold * 100`%.



### Comparison

The `r formatted1_pars$algo` identified `r dim(params$formatted1)[1]` clusters while the `r formatted2_pars$algo` algorithm partitioned the data into the set `r params$formatted2_pars$k` clusters. The largest `r formatted1_pars$algo` cluster had `r params$formatted1$n[[1]]` sequences which represents `r params$formatted1$n_percent[[1]]` of all sequences within the data with the resulting `r params$formatted1_pars$pattern` being `r params$formatted1[1, c(4)]`. The largest `r formatted2_pars$algo` cluster had `r params$formatted2$n[[1]]` sequences which represents `r params$formatted2$n_percent[[1]]` of all sequences within the data with the resulting `r params$formatted2_pars$pattern` being `r params$formatted2[1, c(4)]`.

Running the approxmapR algorithm on the consensus patterns produced by both algorithms
using the following parameters:

* `r formatted3_pars$algo`: k = `r formatted3_pars$k` with the `r formatted3_pars$pattern` filter pattern being used at the threshold of `r formatted3_pars$threshold * 100`%

Resulted in the following:

```{r} 

#::kable(formatted3[, c(1,2,3,4)])
formatted3[, c(1,2,3,4)]

```

All exported files for the `r formatted1_pars$algo` and `r formatted2_pars$algo`, as well as for the comparison can be found `r params$approxmapR_results`.